Testing Rate of Change Transition models

### 2023-05-23 minos/testing/rate_of_change_testing.Rmd
Lots of WIP stuff still on the rate of change testing front. Some sections comparing old (mean regenerative models) with new (rate of change)

out.path <- 'Minos/output/default_config/'
base.dat <- read_singular_local_out(out.path, 'baseline')
```

@@

```{r}
miss.values <- c(-9, -8, -7, -3, -2, -1, -10)
raw <- raw.cv %>%
  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
  group_by(pidp) %>%
  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
  filter(time != 2009)

base <- base.cv %>%
  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
  group_by(pidp) %>%
  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
  filter(time != 2014)
```

```{r}
raw.sub <- raw %>%
  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
  mutate(source = 'raw')
base.sub <- base %>%
  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
  mutate(source = 'output')

combined <- rbind(raw.sub, base.sub)

combined.sum <- combined %>%
  group_by(time, source) %>%
  summarise(hh_income = mean(diff.hh_income, na.rm = TRUE),
            SF12_MCS = mean(diff.SF12_MCS, na.rm = TRUE))

ggplot(combined.sum, aes(x = time, y = hh_income, group = source, color = source)) +
  geom_line()

ggplot(combined.sum, aes(x = time, y = SF12_MCS, group = source, color = source)) +
  geom_line()
```

```{r}
test <- base.cv %>%
  group_by(time) %>%
  summarise(mean.income = mean(hh_income, na.rm = TRUE),
            mean.SF12 = mean(SF_12, na.rm = TRUE))

ggplot(test, aes(x = time, y = mean.income)) +
  geom_line()

ggplot(test, aes(x = time, y = mean.SF12)) +
  geom_line()
```

@@

```{r}
# need baseline and interventions to compare with
out.path <- 'Minos/output/default_config/'
baseline <- read_singular_local_out(out.path, 'baseline')
allchild <- read_singular_local_out(out.path, 'hhIncomeChildUplift')
poverty <- read_singular_local_out(out.path, 'hhIncomePovertyLineChildUplift')
livwage <- read_singular_local_out(out.path, 'livingWageIntervention')
energy <- read_singular_local_out(out.path, 'energyDownlift')
```

```{r}
alchild.plot <- SF12_marg_dist_densigram_plot(baseline,
                              allchild,
                              'All Child Uplift',
                              target.year = 2030)
pov.plot <- SF12_marg_dist_densigram_plot(baseline,
                              poverty,
                              'Poverty Child Uplift',
                              target.year = 2030)
livwage.plot <- SF12_marg_dist_densigram_plot(baseline,
                              livwage,
                              'Living Wage Uplift',
                              target.year = 2030)
energy.p <- SF12_marg_dist_densigram_plot(baseline,
                              energy,
                              'Energy Downlift',
                              target.year = 2030)
```

```{r}
alchild.plot2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
                              allchild,
                              'All Child Uplift',
                              target.year = 2030)
pov.plot2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
                              poverty,
                              'Poverty Child Uplift',
                              target.year = 2030)
livwage.plot2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
                              livwage,
                              'Living Wage Uplift',
                              target.year = 2030)
energy.p2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
                              energy,
                              'Energy Downlift',
                              target.year = 2030)
```

@@

```{r}
income.spag.base <- spaghetti_plot(base.dat, 'hh_income')
income.spag.base
```

This one actually looks worse I think... Might need to have a look at subsetting to see the tight band in the middle in more detail. Also worth using the same plot on raw data to see how that looks:

```{r}
income.spag.raw <- spaghetti_plot(raw.dat, 'hh_income')
income.spag.raw
```

Very useful plot! Much more variability in the raw data that we don't keep hold of in the simulation. Also shows that the spike in 2015 is mostly due to about 5-6 data points by the looks of things. Again worth a chat with the group to figure out where our acceptable thresholds are, as we are unlikely to be able to capture all of the heterogeneity in such a varied and skewed variable like hh_income.

@@

>>>>

# read output and remove dead people
out.path <- 'Minos/output/default_config/'
base.dat <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
```

@@

```{r}
#miss.values <- c(-9, -8, -7, -3, -2, -1, -10)
#raw <- raw.cv %>%
#  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
#  group_by(pidp) %>%
#  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
#         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
#  filter(time != 2009)

#base <- base.cv %>%
#  filter(!hh_income %in% miss.values, !SF_12 %in% miss.values) %>%
#  group_by(pidp) %>%
#  mutate(diff.hh_income = hh_income - lag(hh_income, order_by = time),
#         diff.SF12_MCS = SF_12 - lag(SF_12, order_by = time)) %>%
#  filter(time != 2014)
```

```{r}
#raw.sub <- raw %>%
#  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
#  mutate(source = 'raw')
#base.sub <- base %>%
#  select(pidp, time, diff.hh_income, diff.SF12_MCS) %>%
#  mutate(source = 'output')

#combined <- rbind(raw.sub, base.sub)

#combined.sum <- combined %>%
#  group_by(time, source) %>%
#  summarise(hh_income = mean(diff.hh_income, na.rm = TRUE),
#            SF12_MCS = mean(diff.SF12_MCS, na.rm = TRUE))

#ggplot(combined.sum, aes(x = time, y = hh_income, group = source, color = source)) +
#  geom_line()

#ggplot(combined.sum, aes(x = time, y = SF12_MCS, group = source, color = source)) +
#  geom_line()
```

```{r}
#test <- base.cv %>%
#  group_by(time) %>%
#  summarise(mean.income = mean(hh_income, na.rm = TRUE),
#            mean.SF12 = mean(SF_12, na.rm = TRUE))

#ggplot(test, aes(x = time, y = mean.income)) +
#  geom_line()

#ggplot(test, aes(x = time, y = mean.SF12)) +
#  geom_line()
```

@@

```{r}
# need baseline and interventions to compare with
#out.path <- 'Minos/output/default_config/'
#baseline <- read_singular_local_out(out.path, 'baseline', drop.dead = TRUE)
#allchild <- read_singular_local_out(out.path, 'hhIncomeChildUplift', drop.dead = TRUE)
#poverty <- read_singular_local_out(out.path, 'hhIncomePovertyLineChildUplift', drop.dead = TRUE)
#livwage <- read_singular_local_out(out.path, 'livingWageIntervention', drop.dead = TRUE)
#energy <- read_singular_local_out(out.path, 'energyDownlift', drop.dead = TRUE)
```

```{r}
#allchild.plot <- SF12_marg_dist_densigram_plot(baseline,
#                              allchild,
#                              'All Child Uplift',
#                              target.year = 2030)
#pov.plot <- SF12_marg_dist_densigram_plot(baseline,
#                              poverty,
#                              'Poverty Child Uplift',
#                              target.year = 2030)
#livwage.plot <- SF12_marg_dist_densigram_plot(baseline,
#                              livwage,
#                              'Living Wage Uplift',
#                              target.year = 2030)
#energy.p <- SF12_marg_dist_densigram_plot(baseline,
#                              energy,
#                              'Energy Downlift',
#                              target.year = 2030)
```

```{r}
#alchild.plot2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
#                              allchild,
#                              'All Child Uplift',
#                              target.year = 2030)
#pov.plot2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
#                              poverty,
#                              'Poverty Child Uplift',
#                              target.year = 2030)
#livwage.plot2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
#                              livwage,
#                              'Living Wage Uplift',
#                              target.year = 2030)
#energy.p2 <- SF12_marg_dist_densigram_plot_oneyear(baseline,
#                              energy,
#                              'Energy Downlift',
#                              target.year = 2030)
```

@@

```{r}
income.spag.base <- spaghetti_plot(base.dat, 'hh_income')
income.spag.base

income.spag.raw <- spaghetti_plot(raw.dat, 'hh_income')
income.spag.raw
```

Much more variability in the raw data that we don't keep hold of in the simulation. Also shows that the spike in 2015 is mostly due to about 5-6 data points by the looks of things. Again worth a chat with the group to figure out where our acceptable thresholds are, as we are unlikely to be able to capture all of the heterogeneity in such a varied and skewed variable like hh_income.

#### Comparison

The base spaghetti plot doesn't look very good actually... I think I need a version using the previous models so I'll do a baseline run and save the data somewhere for testing purposes.

```{r}
old.files <- list.files('/home/luke/Documents/WORK/MINOS/PREVIOUS_OUTPUTS/mean_regen_income_sf12/', pattern='[0-9]{4}.csv', full.names = TRUE)
old.dat <- do.call(rbind, lapply(old.files, read.csv))
# remove dead people
old.dat <- old.dat %>%
  filter(alive != 'dead')

income.spag.base <- spaghetti_plot(base.dat, 'hh_income')
income.spag.base

income.spag.old <- spaghetti_plot(old.dat, 'hh_income')
income.spag.old
```

Almost no change. The only noticeable change at this resolution is a small shift up? Where some people who were below the zero line previously seem to have shifted up, and there are some trajectories that have been seemingly raised out of the main group and lie slightly above the average spread. Need to figure out why this is still happening with a completely different set of models...

Very useful plot!

@@

#### Comparison

sf12.spag.old <- spaghetti_plot(old.dat, 'SF_12')
sf12.spag.old
```
